---
title: "Survey Data"
author: "Trevor Johnson"
date: "11/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Libraries

```{r}
library(data.table)
library(tidyverse)
library(janitor)
```

Read in data

```{r}
# Original data file isn't saved to github, only google drive.
# Download the survey data and the questions_scores.csv file.
# Then specify the path where you saved the files
# After that, everything else below should work
path <- "~/Desktop/data"

df <- "W241_Survey_Results.csv" %>% file.path(path, .) %>% fread
q_scores <- "question_scores.csv" %>% file.path(path, .) %>% fread
```

Set up mappings for later

```{r}
agree_mapping <- data.table(
  desc = c("Disagree", "Somewhat Disagree", "Neutral", "Somewhat Agree", "Agree"),
  score = 1:5
)

letter_mapping <- data.table(
  letter = c("E", "A", "C", "N", "O"), 
  personality = c("Extroversion", "Agreeableness", "Conscientiousness", "Neuroticism", "Openness to Experience")
)
```


Get question names. Don't need to re-run this. This is just to extract the questions for google sheets

```{r}
# extract the question portion of the title
# questions <- names(df)[grepl("Please indicate the extent", names(df))] %>% 
#   str_replace("Please indicate the extent to which you agree or disagree with each statement. - ", "")

# copy questions to clipboard so we can paste them into google sheets for score lookups
# clip <- pipe("pbcopy", "w")
# write.table(data.frame(questions), file=clip,row.names=FALSE)    
# close(clip)
```

Data cleaning

```{r}
# make the first row be the column names because they are more descriptive
names(df) <- unlist(df[1,])

# first row can be removed for actual analysis
df <- df[2:nrow(df),]

# Clean up question columns
old_names <- names(df)[grepl("Please indicate the extent", names(df))]
new_names <- old_names %>% str_replace("Please indicate the extent to which you agree or disagree with each statement. - ", "")
setnames(x = df, old = old_names, new = new_names)

# more cleaning
df <- df %>% 
  
  # this function cleans up column names to make all spaces underscores, and removes special chars  
  janitor::clean_names() %>% 

  # fix some date columns
  mutate_at(c("start_date", "end_date", "recorded_date"), list(~as.Date(., "%Y-%m-%d"))) %>% 

  # rename a few cols w long names
  rename(
  dob_month = please_enter_your_date_of_birth_month_please_select,
  dob_day = please_enter_your_date_of_birth_day_please_select, 
  education_highest = what_is_the_highest_level_of_school_you_have_completed_or_the_highest_degree_you_have_received,
  familiar_with_test = are_you_familiar_with_the_five_factor_personality_test,
  taken_test_before = have_you_taken_the_five_factor_personality_test_before,
  compliance_q_glass = according_to_the_video_you_just_watched_what_style_of_glass_pitcher_was_produced,
  compliance_q_mclaughlin = which_of_the_following_animals_was_not_shown_in_the_video_you_just_watched)


# get question names again to use later
q_cols <- df %>% select(i_am_the_life_of_the_party:i_am_full_of_ideas) %>% names()

# convert to long format
df_long <- melt.data.table(df, measure.vars = q_cols, variable.name = "question", value.name = "answer") %>% 
  
  # join w/ the score mappings
  left_join(select(q_scores, q_clean, letter, plus_minus), by = c("question" = "q_clean")) %>% 
  
  # convert agree/disagree etc. to numeric
  left_join(agree_mapping, by = c("answer" = "desc")) %>% 
  mutate(score2 = score * plus_minus)


# save the full dataset
# fwrite(df, "full_data.csv")
```

Check out fishy responses

```{r}
# 6 people answered neutral all the way down
df_long[,.N, by = c("response_id", "answer")][answer=="Neutral"]

# 10 people just put the same thing every time
df_long[,.N, by = c("response_id", "answer")][N == 50]

# 17 people basically put the same answer all the way down, with only slight variations
df_long[,.N, by = c("response_id", "answer")][N >= 45]
```

Create function to compute score for a user given response_id.

```{r}
calc_score <- function(id, df = df_long, wide_output=FALSE){
  score_df <- df[response_id == id][,.(score = sum(score2)), by = "letter"]
# each personality type has a baseline per the documentation
  score_df[, baseline := 
             ifelse(letter == "E", 20, 
             ifelse(letter == "A", 14, 
             ifelse(letter == "C", 14, 
             ifelse(letter == "N", 38, 
             ifelse(letter == "O", 8, 
                    # use 999 as an error
                    99999)))))]
  
  score_df[, final_score := baseline + score]
  score_df <- score_df[, .(letter, final_score)]
  
  if (wide_output){
    
    df_wide <- data.frame(matrix(ncol=5))
    df_wide[1,] <- score_df$final_score
    names(df_wide) <- c("personality_e", "personality_a", 
                        "personality_c", "personality_n", "personality_o")
    df_wide$response_id <- id
    
    return(df_wide)
    
  }else{
    return(data.frame(score_df))
  }
  
}

# test it out
calc_score("R_10xBdoxbP4TBlzD")
calc_score("R_10xBdoxbP4TBlzD", wide_output = TRUE)
```

Aggregate everyone's results

```{r}
agg_results <- function(ids, df = df_long){
  
  # initialize df w/ 1 person's score
  agg_results_df <- calc_score(ids[1], df_long)

  # loop through all unique ids calculating their scores and stack the df's
  for (i in 2:length(ids)){
    agg_results_df <- rbind(agg_results_df, calc_score(ids[i], df_long))
  }
  
  # convert to data.table
  setDT(agg_results_df)
  
  # average score by letter overall. Then join w/ letter mapping to get personality trait names.
  agg_results_df <- agg_results_df[, .(avg_score = round(mean(final_score)),1), by = letter] %>% 
    left_join(letter_mapping, by = "letter") %>% 
    select(personality, avg_score)

  return(agg_results_df)
}

# get all unique ids
ids <- df_long$response_id %>% unique()

# run function to aggregate all ids
agg_results_df <- agg_results(ids, df_long)
```

Make a bar chart of avg scores

```{r}
agg_results_df <- agg_results_df %>% arrange(personality)
ggplot(agg_results_df, aes(x = personality, y = avg_score)) + 
  geom_col(fill = c(rgb(0/255, 50/255, 98/255))) + 
  annotate("text", x = 1:5, y = agg_results_df$avg_score-3, 
           label = agg_results_df$avg_score, color = "white") +
  #coord_flip() +
  labs(title = "Aggregated Personality Trait Scores", x = "", y = "Average Score") +
  theme_minimal()
```


```{r}
df2 <- df %>% 
  
  # this function cleans up column names to make all spaces underscores, and removes special chars  
  janitor::clean_names() %>% 

  # fix some date columns
  #mutate_at(c("start_date", "end_date", "recorded_date"), funs(as.Date(., "%m/%d/%Y")))
  mutate_at(c("start_date", "end_date", "recorded_date"), list(~as.Date(., "%m/%d/%Y")))


df2 %>% select(contains("_date")) %>% glimpse

```


Get scores for each person, and create a clean dataset


```{r}
all_scores_df <- calc_score(ids[1], wide_output = TRUE)
for (i in 2:length(ids)){
  all_scores_df <- rbind(all_scores_df, calc_score(ids[i], wide_output = TRUE))
}

setDT(all_scores_df)
```

Join with big dataset and perform final clean

```{r}
df2 <- df %>% select(-i_am_the_life_of_the_party:-i_am_full_of_ideas) %>% 
  left_join(all_scores_df, by = "response_id")

# variables to convert to numeric
num_vars <- c("progress", "duration_in_seconds", "dob_day", "timing_first_click", "timing_last_click",
              "timing_page_submit", "timing_click_count", "timing_first_click_2", "timing_last_click_2",
              "timing_page_submit_2", "timing_click_count_2", "age", "gender", "hhi", "ethnicity", 
              "hispanic", "education", "political_party", "region")

df2[, (num_vars) := lapply(.SD, as.numeric), .SDcols=num_vars]
```

Save file

```{r}
#fwrite(df2, "~/Desktop/W241_Survey_Results_Clean.csv")
```

Do some analysis with charts

```{r}
df3 <- df2 %>% 
  mutate(treatment_type = 
           ifelse(compliance_q_mclaughlin == "Fish", "Mclaughlin", 
           ifelse(compliance_q_glass == "Confetti (colored)", "Glass blowing", "Non complier"))) %>% 
  select(personality_e:treatment_type) %>% 
  melt.data.table(id.vars = "treatment_type", variable.name = "personality", value.name = "score") %>% 
  mutate(personality = toupper(substr(personality, 13, 13))) %>% 
  rename(letter = personality) %>% 
  left_join(letter_mapping, by = "letter")


df3_agg <- df3[, .(score = mean(score)), by = c("treatment_type", "personality")]

df3_agg %>% 
  ggplot(aes(x = personality, y = score, fill = treatment_type)) + 
  geom_col(position = "dodge") + 
  theme_minimal()


df3[, .(.N / 5), by = c("treatment_type")]


agg_results_df <- agg_results_df %>% arrange(personality)
ggplot(agg_results_df, aes(x = personality, y = avg_score)) + 
  geom_col(fill = c(rgb(0/255, 50/255, 98/255))) + 
  annotate("text", x = 1:5, y = agg_results_df$avg_score-3, 
           label = agg_results_df$avg_score, color = "white") +
  #coord_flip() +
  labs(title = "Aggregated Personality Trait Scores", x = "", y = "Average Score") +
  theme_minimal()
```







